### 2019-09-08 Writeup/Thesis/chapters/abstract.tex
Changes; - Final check for gramatical errors
Changes; - Small corrections added

Many encrypted connections require the comparison of a fingerprint to protect against eavesdropping. A substantial amount of past work has aimed to propose fingerprint representations that work better with human limitations. "Trustwords" proposed by pEp is an example of such a scheme, where the fingerprint is encoded as words in an attempt to improve usability.

This works main aim is to asses if Trustword's recommend minimum number of four words is sufficient.
In order to achieve this goal, this work implements an attack on Trustwords and quantifies its effectiveness on more than 400 participants.  A tool called GreenOnion was designed to assist in quantifying attack feasibility. GreenOnion improved substantially on a similar tool's ability to search for matches concurrently where it was able to search for over 1.5 million concurrent keys. Our findings show a substantial increase in attack success compared to related literature. We believe this increase is due to levering the design flaws in the Trustword scheme. Therefore, we believe a minimum of four Trustwords is insufficient to provide even a basic level of security.

>>>>

Many encrypted connections require the comparison of a fingerprint to protect against eavesdropping. A substantial amount of past work has aimed to propose fingerprint representations that work better with human limitations. "Trustwords" proposed by p$\equiv$p is an example of such a scheme, where the fingerprint is encoded as words in an attempt to improve usability.

This works main aim is to assess if Trustword's recommended minimum number of four words is sufficient.
In order to achieve this goal, this work implements an attack on Trustwords and quantifies its effectiveness on more than 400 participants.  A tool called GreenOnion was designed to assist in quantifying attack feasibility. GreenOnion improved substantially on a similar tool's ability to search for matches concurrently. Our findings show a substantial increase in attack success compared to related literature. We believe this increase is due to levering the design flaws in the Trustword scheme. Therefore, we believe a minimum of four Trustwords is insufficient to provide even a basic level of security.

### 2019-09-08 Writeup/Thesis/chapters/design.tex
Changes; - Small corrections added

\subsection{Soundex}
Soundex is one of the most famous example of a phonetic algorithm (See Section \ref{sec:soundex}). This is due in part to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}.

Soundex, however, has had extensive work focused on its own deficiencies alongside quantification of its performance under certain conditions. However, even with these previously discussed issues, its algorithmic simplicity and popularity allows it to remain relevant for assessment in this project.

@@

\subsection{Metaphone}
Metaphone, as discussed in Section \ref{sec:metaphone} is another famous phonetic algorithm. Metaphone is arguably on the same level of ubiquity as that of Soundex with it finding itself implemented in languages such as PHP\cite{php}. Metaphone was chosen due to similar reasons to that of Soundex and NYSIIS.

@@

These attacks are ascending in complexity. The timeframe for computing a near-collision was decided as one week with attack strengths of 7 GPU-days \textit{(1 x GPU)}, 70 GPU-days \textit{(10 x GPUs)} and 700 GPU-days \textit{(100 x GPUs)} on a mid-range GPU (AMD RX 480). Due to the long lifetime of PGP keys, 7 days was chosen as a reasonable length of the attack. The tool requires no user interaction once started, therefore, it can be left for an extended period to compute potential keys. This timeframe is also a maximum time, as these attacks also include keys that have a shorter average computation time.

>>>>

\subsection{Soundex}
Soundex is one of the most famous examples of a phonetic algorithm (See Section \ref{sec:soundex}). This is due in part to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}.

Soundex, however, has had extensive work focused on its deficiencies alongside quantification of its performance under certain conditions. However, even with these previously discussed issues, its algorithmic simplicity and popularity allows it to remain relevant for assessment in this project.

@@

\subsection{Metaphone}
Metaphone, as discussed in Section \ref{sec:metaphone}, is another famous phonetic algorithm. Metaphone is arguably on the same level of ubiquity as that of Soundex with it finding itself implemented in languages such as PHP\cite{php}. Metaphone was chosen due to similar reasons to that of Soundex and NYSIIS.

@@

These attacks are ascending in complexity. The timeframe for computing a near-collision was decided as one week with attack strengths of 7 GPU-days \textit{(1 x GPU)}, 70 GPU-days \textit{(10 x GPUs)} and 700 GPU-days \textit{(100 x GPUs)} on a mid-range GPU (AMD RX 480). Due to the long lifetime of PGP keys, 7 days was chosen as a reasonable length of the attack. The tool requires no user interaction once started. Therefore, the tool can be left for an extended period to compute potential keys. This timeframe is also a maximum time, as these attacks also include keys that have a shorter average computation time.

### 2019-09-08 Writeup/Thesis/chapters/evaluation.tex
Changes; - Small corrections added

Research Question 1 asks: \textit{“What are the best performing schemes to quantify phonetic similarity?”}. This paper has discussed several metrics used to quantify phonetic similarity (See Section \ref{sec:metrics}). Most metrics were chosen due to their high utilization and occurrence in literature. The only exception to this is the presence of Phonetic Vectors. Performance of these metrics were then quantified in a direct study that selected the top three from a set of five (See Section \ref{exp:metric}). These newly selected metrics was then assessed in an attack context with their performance being indirectly quantified (See Section \ref{sec:exp2_design}).

@@

\section{Research Question \ref{goal:complexity}}
Research Question \ref{goal:complexity}: \textit{``What is the time and computation complexity required to generate a ’similar’ keys for a targeted key pair?''} has been discussed by the project alongside the actual generation of keys as shown in Section \ref{sec:key_gen}. Furthermore, the distribution of key permutations on a set of real-world keys were discussed in Section \ref{sec:averagePerms} and \ref{sec:vulnKeys}. To achieve this, the tool GreenOnion was developed. Inspiration for GreenOnion was gathered from a currently available tool known as Scallion. Experimentation results were gathered and displayed in Section \ref{sec:SvG} that showed the significant increase in the number of near-collision keys GreenOnion can concurrently search for. GreenOnion is able to handle in excess of 1.5 million keys. This capability is in comparison to the maximum of 2513 keys the alternative; Scallion can compare. We believe this is due to the implementation of the bloom filter. Its unique characteristics of $O(k)$ complexity regardless of the number of elements in the structure allows a huge number of concurrent checks to occur.

@@

\subsection*{Trustword improvements}
The first area of proposed work could be recommendations into how to improve Trustwords backed by empirical evidence. We feel the most promising avenue would be the utilisation of Phonetic Vectors unique qualities to identify words of low quality by measuring very low vector distances. For example, present in the dictionary are the words \verb|THERE| and \verb|THEIR|, these could be massively improved upon and could result in a better quality word list. Moreover, utilising the quantification of dissimilarity could be used to create a dictionary of maximized phonetic distance. This phonetically distinct wordlist could then be assessed in a similar way with its success rate quantified against users.

@@

\subsection*{GreenOnion optimisations}
A minimal amount of work was allocated to improving the performance of GreenOnion as low-level optimizations are very time-consuming. A project, therefore, could aim to improve on the performance already recorded in this paper. This project could improve the performance of the attack and allow for more effective keys to be computed in less time.

>>>>

Research Question 1 asks: \textit{“What are the best performing schemes to quantify phonetic similarity?”}. This paper has discussed several metrics used to quantify phonetic similarity (See Section \ref{sec:metrics}). Most metrics were chosen due to their high utilisation and occurrence in literature. The only exception to this is the presence of Phonetic Vectors. Performance of these metrics were then quantified in a direct study that selected the top three from a set of five (See Section \ref{exp:metric}). These newly selected metrics was then assessed in an attack context with their performance being indirectly quantified (See Section \ref{sec:exp2_design}).

@@

\section{Research Question \ref{goal:complexity}}
Research Question \ref{goal:complexity}: \textit{``What is the time and computation complexity required to generate a ’similar’ keys for a targeted key pair?''} has been discussed by the project alongside the actual generation of keys as shown in Section \ref{sec:key_gen}. Furthermore, the distribution of key permutations on a set of real-world keys were discussed in Section \ref{sec:averagePerms} and \ref{sec:vulnKeys}. To achieve this, the tool GreenOnion was developed. Inspiration for GreenOnion was gathered from a currently available tool known as Scallion. Experimentation results were gathered and displayed in Section \ref{sec:SvG} that showed the significant increase in the number of near-collision keys GreenOnion can concurrently search for. GreenOnion is able to handle in excess of 1.5 million keys. This capability is in comparison to the maximum of 2513 keys the alternative; Scallion can compare. We believe this is due to the implementation of the bloom filter. Its unique characteristics of $O(k)$ complexity, regardless of the number of elements in the structure allows a huge number of concurrent checks to occur.

@@

\subsection*{Trustword improvements}
The first area of proposed work could be recommendations into how to improve Trustwords backed by empirical evidence. We feel the most promising avenue would be the utilisation of Phonetic Vectors unique qualities to identify words of low quality by measuring very low vector distances. For example, present in the dictionary are the words \verb|THERE| and \verb|THEIR|, these could be massively improved upon and could result in a better quality word list. Moreover, utilising the quantification of dissimilarity could be used to create a dictionary of maximised phonetic distance. This phonetically distinct wordlist could then be assessed in a similar way with its success rate quantified against users.

@@

\subsection*{GreenOnion optimisations}
A minimal amount of work was allocated to improving the performance of GreenOnion as low-level optimisations are very time-consuming. A project, therefore, could aim to improve on the performance already recorded in this paper. This project could improve the performance of the attack and allow for more effective keys to be computed in less time.

### 2019-09-08 Writeup/Thesis/chapters/implementation.tex
Changes; - Small corrections added

\section{MainExperiment}
In order to assess users in a similar scenario to ones experienced when utilizing \pep in the real-world, a custom application was required. The front-end is developed in Javascript with a Python-Flask backend that acts as the functionality of the webpage. Several endpoints are exposed like \verb|get_audio| or \verb|get_words| that are requested and displayed on the front-end.

>>>>

\section{MainExperiment}
In order to assess users in a similar scenario to ones experienced when utilising \pep in the real-world, a custom application was required. The front-end is developed in Javascript with a Python-Flask backend that acts as the functionality of the webpage. Several endpoints are exposed like \verb|get_audio| or \verb|get_words| that are requested and displayed on the front-end.

### 2019-09-08 Writeup/Thesis/chapters/introduction.tex
Changes; - More wording fixes
Changes; - Final check for gramatical errors
Changes; - Small corrections added

Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating a more secure and user-friendly fingerprint encoding remains an important task.

One proposed scheme claiming increased usability is Pretty Easy Privacy's (p$\equiv$p) "Trustwords". Trustwords is an implementation of word fingerprint mapping. This claimed increase in usability is achieved by the user comparing a reduced number of words compared to alternatives. This potential usability boost, however, comes at a cost of a much larger word list.

@@

\newpage
The report is structured as follows:

@@

    \item \textbf{Chapter 5} discusses the significant implementation details of the designs formalized in Chapter 4.

>>>>

Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating secure and user-friendly fingerprint encodings remains an important task.

One proposed scheme claiming increased usability is Pretty Easy Privacy's (p$\equiv$p) "Trustwords". Trustwords is an implementation of word fingerprint mapping. This claimed increase in usability is achieved by the user comparing a reduced number of words compared to alternatives. This potential usability boost, however, comes at the cost of a much larger word list.

@@

The report is structured as follows:

@@

    \item \textbf{Chapter 5} discusses the significant implementation details of the designs formalised in Chapter 4.

### 2019-09-08 Writeup/Thesis/chapters/lit_review.tex
Changes; - Small corrections added

\textbf{OpenSSH}. OpenSSH Visual Host Key is an ASCII based graphical scheme used in MiTM detection when connecting to SSH servers. The image is generated by the movement of the central marker across a grid. Every 2-bits of the generated MD5 hash defines the diagonal direction of movement. For example, if the first 2-bits are \verb|11|, the marker moves South East. The different characters are determined by the number of times the marker is present in that index, for example, if the maker is present on a square 4 times the character is an equals sign (\verb|=|).

@@

Tables \ref{tab:times} \& \ref{tab:results} contain the accuracy results of all papers assessed; this is to aid in visual comparison. Each paper used a different metric for measuring accuracy. Therefore, all results have been translated into ``overall accuracy.''. For example, if a paper presented "security failures" or "attack success rate" of 5\%, the overall accuracy of the scheme is 95\%.

@@

Research has assessed the performance of these schemes and how they affect the security of the authentication ceremony. The literature agrees on CaC being the best overall scheme to compare fingerprints \cite{tan2017can}\cite{uzun2007usability} with CaS highlighted for its poor security and usability. CaE has had conflicting results. \cite{uzun2007usability} discarded it after one round due to ``poor usability.'' However, \cite{tan2017can} considered it the best method overall for usability and security. These conflicting results, therefore, show polarisation in the results of CaE. However, this could be due to the different overall use-cases of the studies. Validation of results from either study would, therefore, be an area of additional study.

@@

\subsubsection*{Conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilisation of encoding schemes in realistic conditions, i.e. ``remove vs proximity." This topic was initially covered by \cite{shirvanian2017pitfalls}, but their scope was limited.

@@

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests, the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. MadLibs\footnote{\url{https://en.wikipedia.org/wiki/Mad\_Libs}} was the base for these sentences. MadLibs is where static placeholders are replaced with a member from a list of potential words.\\
This work is an extension to work performed by \textbf{Juola} and \textbf{Zimmermann}\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's findings are limited by the lack of empirical data backing up claims made by the author the performance of the system, and security is only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings -- The first being by \textbf{Michael Rogers}\footnote{\url{https://github.com/akwizgran/basic-english}}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical}. Earlier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.
Aside from this research, there have been further informal implementations of fingerprint encodings  The first being by \textbf{Michael Rogers}\footnote{\url{https://github.com/akwizgran/basic-english}}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical}. Earlier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

\newpage
\subsubsection*{Pretty Easy Privacy}
\label{sec:pep}
A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\footnote{\url{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The embedded design principles state that above all the systems should be easy to install, use and understand.

>>>>

\textbf{OpenSSH}. OpenSSH Visual Host Key is an ASCII based graphical scheme used in MiTM detection when connecting to SSH servers. The image is generated by the movement of the central marker across a grid. Every 2-bits of the generated MD5 hash defines the diagonal direction of movement. For example, if the first 2-bits are \verb|11|, the marker moves South East. The different characters are determined by the number of times the marker is present in that index, for example, if the marker is present on a square 4 times the character is an equals sign (\verb|=|).

@@

Tables \ref{tab:times} \& \ref{tab:results} contain the accuracy results of all papers assessed; this is to aid in visual comparison. Each paper used a different metric for measuring accuracy. Therefore, all results have been translated into ``overall accuracy.'' For example, if a paper presented ``security failures'' or ``attack success rate'' of 5\%, the overall accuracy of the scheme is 95\%.

@@

Research has assessed the performance of these schemes and how they affect the security of the authentication ceremony. The literature agrees on CaC being the best overall scheme to compare fingerprints \cite{tan2017can}\cite{uzun2007usability} with CaS highlighted for its poor security and usability. CaE has had conflicting results. \cite{uzun2007usability} discarded it after one round due to ``poor usability.'' However, \cite{tan2017can} considered it the best method overall for usability and security. These conflicting results, therefore, show polarisation in the results of CaE. However, this could be due to the different overall use-cases of the studies. Validation of results from either study would, therefore, be an area of additional work.

@@

\subsubsection*{Conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilisation of encoding schemes in realistic conditions, i.e. ``remote vs proximity." This topic was initially covered by \cite{shirvanian2017pitfalls}, but their scope was limited.

@@

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests, the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. MadLibs\footnote{\url{https://en.wikipedia.org/wiki/Mad\_Libs}} was the base for these sentences. MadLibs is where static placeholders are replaced with a member from a list of potential words.\\
This work is an extension to work performed by \textbf{Juola} and \textbf{Zimmermann}\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's findings are limited by the lack of empirical data backing up claims made by the author on the performance of the system, and security is only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings -- The first being by \textbf{Michael Rogers}\footnote{\url{https://github.com/akwizgran/basic-english}}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical}. Earlier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

\subsubsection*{Pretty Easy Privacy}
\label{sec:pep}
A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\footnote{\url{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The embedded design principles state that above all the systems should be easy to install, use and understand.
